\documentclass{article}
\begin{document}
Much has been said about how the rapid reduction in cost of the semiconductor has, in the last 60 years, changed the world in a significant way. The rapid spread of inexpensive and energy efficient computers, networks, and storage facilities, has revolutionised how we access information, exchange goods and services, and communicate with one another. The diffusion of the smartphone, specifically, has brought forth an explosion in the amount of information generated globally, with more than ?? billions users who spend ?? of their days interacting with their phones, browsing the web, using applications, generating tracking data, and chatting. The latter usage in particular is one of the most popular. The top downloaded applications among most online distributors are ??messenger/whatsapp/wechat??, with more than a billion active user each month. People spend on average ?? within a chat interface.
Besides keeping up with friends and professional contacts, business transactions are also conducted through chat, either arranging sale and delivery of goods, or for customer assistance, the former being much more prevalent in Asia, where most medium and large size companies, as well as some smaller ones, have a WeChat presence (conversational commerce). Increasingly, many of these transactions are being automated through the deployment of chatbots (bots). The most popular bot platform is Facebook messenger, which introduced the functionality to developers in April 2016??, and it has since taken off with more than ?? bots on the platform. The development of Voice Assistants such as Google Assistant, Siri, Cortana, Alexa, and open-sourced Mycroft, have also pushed the deployment of conversational interfaces, and to some extent through the chat medium, with all the mentioned agents providing some form of textual input.
Besides the marketing pushes, the use of chatbots has increased thanks to advances in Natural Language Processing (NLP) and Natural Language Understanding (NLU)
tools available to developers \cite{JavierCouto}.
While early chatbot implementations relied on simple pattern matching rules based on recognition of specific words (entity recognition) or parts-of-speech (POS), most of today's chatbot frameworks can leverage large corpora to apply machine learning algorithm, such as Intent analysis. Conversation can follow a slot or flow model: the latter is a hardcoded scripted flow diagram that guides the user through a preset conversation; the former specifies 'slots' that contain some data the developer is interested in, and the chatbot will use NLP techniques to fill the slots from conversations with the user. Responses are typically pre-written and matched to an intent, but advances in deep learning are opening up possibilities for generative models, which create the answer from scratch\cite{Gregori}. \\

TODO: The advantages of chatbot interfaces over traditional ones and over communicating.
People perceive email to be a ``faster'' way to communicate than email
Sentiment analysis

Studies in medical settings (and social characteristics of chatbots)


nutrition motivation for quantified self

nutritional chat

Image recognition
\end{document}
